---
title: "R and RStudio: Getting Started"
---

## Installing R and RStudio

### Steps to install R:
- Go to [r-project.org](https://www.r-project.org/) and download R.

### Installing RStudio:
- Go to [rstudio.com](https://www.rstudio.com/) and download the free version of RStudio.

## Getting Started with RStudio

### Opening RStudio: Interface Overview
- **Main components**: Script editor, console, environment, and plots panes.

### Creating a new R Script:
1. Go to **File** > **New File** > **R Script**.
2. Save your script: Always save your work as an `.R` file.

## Basic R Operations

### Example: Simple Math Operations in R

```{r}
2 + 2
10 / 5
```

Functions:
mean(): Calculate the average.
summary(): Quick data summary.


## Loading Datasets in R

Using built-in datasets:
```{r}
data(mtcars)
```


## Basic Plotting in R

Simple scatter plot:
```{r}
plot(mtcars$mpg)
```
Customizing the plot:
Add labels and title: plot(mtcars$mpg, main="MPG Plot", xlab="Car Index", ylab="MPG").
Explore the dataset: head(mtcars), summary(mtcars).
What’s inside the dataset? Overview of rows and columns.


## Creating and Managing Projects in RStudio

Why create a project?
Keeps files organized for each analysis.
Creating a new project:
File > New Project > New Directory.
Save scripts, data, and results in one folder for easy access.
